---
title: "06b_sceG&T_seuratPCA"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Setup of the workenvironment

## Set root directory for the rest of our analysis

```{r "setup", include=FALSE}
### Ignore code, this is used to set the workig directory. 
rm(list=ls())
t=getwd()
tt=gsub('.{30}$', '', t)
require("knitr")
opts_knit$set(root.dir = tt)
rm(t,tt)
```

## Load necessary libraries for the analysis

```{r}
library(Seurat)
library(dplyr)
```

## Parameters to specify

```{r}
Colourblind_palette <- c( "#E69F00" , "#56B4E9" , "#009E73" , "#F0E442" , "#0072B2" , "#D55E00" , "#CC79A7" , "#000000" )
```

## load Seurat object

```{r}
load(file="Data/RNA/MEL006/SCEsets/Seurat_methodTimepoint.RData")
```

## PCA on the total dataset

next we perform PCA on the scaled data. By default, the genes in object@var.genes are used as input, but can be alternatively defined. Running dimensionality reduction on highly variable genes can improve performance.

```{r}
bc.integrated <- RunPCA(bc.integrated, features = VariableFeatures(object = bc.integrated), seed.use = 13)
```

```{r}
print(bc.integrated[["pca"]], dims = 1:5, nfeatures = 5)
```

## Vizualization of the PCA results

### Plot of PC1 and PC2 coloured by Sample origin

```{bash}
cd Results/MEL006/RNA
mkdir -p Seurat/Method_timepoint/PCA
```

```{r}
PCAPlot(
  object = bc.integrated,
  dims = c(2, 1),
  cols = Colourblind_palette,
  pt.size = 1.5,
  group.by = "method_timepoint")
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/PCA/PCA_1_2.png" , width = 10 , height = 5 )
```

### Plot of PC1 and PC3 coloured by Sample origin

```{r}
PCAPlot(
  object = bc.integrated,
  dims = c(3, 1),
  cols = Colourblind_palette,
  pt.size = 1.5,
  group.by = "method_timepoint")
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/PCA/PCA_1_3.png" , width = 10 , height = 5 )
```

### Plot of PC1 and PC2 coloured by number of expressed genes

```{r}
FeaturePlot(
  object = bc.integrated,
  features = c("nFeature_RNA"),
  dims = c(2, 1),
  pt.size = 1.5,
  reduction = "pca")
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/PCA/PCA_1_2_exprGenes.png" , width = 10 , height = 5 )
```

### Plot of PC1 and PC2 coloured by number of expressed genes

```{r}
FeaturePlot(
  object = bc.integrated,
  features = c("nFeature_RNA"),
  dims = c(3, 1),
  pt.size = 1.5,
  reduction = "pca")
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/PCA/PCA_1_3_exprGenes.png" , width = 10 , height = 5 )
```

## Correlation of number of expressed genes with principal components

```{r}
PCs_corrExpressedGenes <- vector(length = 10)
for (i in 1:10) {
  PCs_corrExpressedGenes[i] <- cor(bc.integrated@reductions$pca@cell.embeddings[,i] , bc.integrated@meta.data$nFeature_RNA )
  names(PCs_corrExpressedGenes)[i] <- paste0("PC",i)
}
PCs_corrExpressedGenes
```

### Visualize top genes associated with principal components

### Dotplot

```{r}
VizDimLoadings(bc.integrated, dims = 1:16, reduction = "pca" , ncol = 8, nfeatures = 50)
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/PCA/PCAdotplot_PC1-16.png" , width = 30 , height = 15 )
```

#### Heatmap

Draws a heatmap focusing on a principal component. Both cells and genes are sorted by their principal component scores. Allows for nice visualization of sources of heterogeneity in the dataset.

```{r}
DimHeatmap(bc.integrated, dims = 1:16, cells = 500, balanced = TRUE, nfeatures = 50 , ncol = 8, fast = FALSE)
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/PCA/PCAheatmap_PC1-16.png" , width = 30 , height = 15 )
```

## Save Seurat Object

```{r}
save(bc.integrated,file="Data/RNA/MEL006/SCEsets/Seurat_methodTimepoint.RData")
```

## Sesion information

```{r}
sessionInfo()
```

