---
title: "09b_sceG&T_seuratTSNE"
author: "Sebastiaan Vanuytven"
date: "28/08/2019"
output: html_document
---

```{r "setup", include=FALSE}
### Ignore code, this is used to set the workig directory. 
rm(list=ls())
t=getwd()
tt=gsub('.{30}$', '', t)
require("knitr")
opts_knit$set(root.dir = tt)
rm(t,tt)
```

## Load necessary libraries for the analysis

```{r}
library(Seurat)
library(dplyr)
library(viridis)
```

## Parameters to specify

```{r}
use.pcs = 1:7
Colourblind_palette <- c( "#E69F00" , "#56B4E9" , "#009E73" , "#F0E442" , "#0072B2" , "#D55E00" , "#CC79A7" , "#000000" )
```

## load Seurat object

```{r}
load(file="Data/RNA/MEL006/SCEsets/Seurat_methodTimepoint.RData")
```

## Run tSNE

tSNE dimensionality reduction plots are then used to visualise clustering results. As input to the tSNE, you should use the same PCs as input to the clustering analysis.

```{r}
bc.integrated <- RunTSNE(
  object = bc.integrated,
  dims = use.pcs,
  do.fast = TRUE)
```

## Vizualize tSNE

```{bash}
cd Results/MEL006/RNA
mkdir -p Seurat/Method_timepoint/tSNE
```

### Standard tSNE plot coloured by cluster

```{r}
TSNEPlot(object = bc.integrated, pt.size=3 , label = T , label.size = 12)
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/tSNE/tSNE_standCol.png" , width = 10 , height = 5 )

TSNEPlot(object = bc.integrated, pt.size=3 , cols = viridis(length(unique(bc.integrated@active.ident))) , label = T , label.size = 12)
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/tSNE/tSNE_ViridisCol.png" , width = 10 , height = 5 )
```

### Standard tSNE plot coloured by Sample origin

```{r}
TSNEPlot(object = bc.integrated, pt.size=3 , cols = Colourblind_palette , group.by = "method_timepoint")
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/tSNE/tSNE_SampleOrigin.png" , width = 10 , height = 5 )
```

### Standard tSNE plot coloured by cluster and shaped by sample origin

```{r}
TSNEPlot(object = bc.integrated, pt.size=3 , label = T , group.by = "method_timepoint", label.size = 12)
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/tSNE/tSNE_standCol_shapeSource.png" , width = 10 , height = 5 )

TSNEPlot(object = bc.integrated, pt.size=3 , cols = viridis(length(unique(bc.integrated@active.ident))) , label = T , group.by = "method_timepoint", label.size = 12)
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/tSNE/tSNE_ViridisCol_shapeSource.png" , width = 10 , height = 5 )
```

## TSNE plots coloured by each score of PC

```{r}
FeaturePlot(bc.integrated, features = c("PC_1", "PC_2", "PC_3", "PC_4", "PC_5", "PC_6", "PC_7", "PC_8", "PC_9", "PC_10") , combine = T, reduction = "tsne")
ggsave( "Results/MEL006/RNA/Seurat/Method_timepoint/tSNE/tSNE_PCscores.png" , width = 20 , height = 10 )
```

## Save Seurat Object

```{r}
save(bc.integrated,file="Data/RNA/MEL006/SCEsets/Seurat_methodTimepoint.RData")
```


## Sesion information

```{r}
sessionInfo()
```


