---
title: "08b_sceG&T_seuratClusterIdentification"
author: "Sebastiaan Vanuytven"
date: "28/08/2019"
output: html_document
---

```{r "setup", include=FALSE}
### Ignore code, this is used to set the workig directory. 
rm(list=ls())
t=getwd()
tt=gsub('.{30}$', '', t)
require("knitr")
opts_knit$set(root.dir = tt)
rm(t,tt)
```

## Load necessary libraries for the analysis

```{r}
library(Seurat)
library(dplyr)
```

## Parameters to specify

```{r}
use.pcs = 1:7
Colourblind_palette <- c( "#E69F00" , "#56B4E9" , "#009E73" , "#F0E442" , "#0072B2" , "#D55E00" , "#CC79A7" , "#000000" )
```

## load Seurat object

```{r}
load(file="Data/GandT/RNA/SCEsets/Seurat.RData")
```

## Cluster Identification

Seurat implements an graph-based clustering approach. Distances between the cells are calculated based on previously identified PCs. Seurat approach was heavily inspired by recent manuscripts which applied graph-based clustering approaches to scRNAseq data. Briefly, Seurat identify clusters of cells by a shared nearest neighbor (SNN) modularity optimization based clustering algorithm. First calculate k-nearest neighbors (KNN) and construct the SNN graph. Then optimize the modularity function to determine clusters.

The FindClusters function implements the procedure, and contains a resolution parameter that sets the granularity of the downstream clustering, with increased values leading to a greater number of clusters. I tend to like to perform a series of resolutions, investigate and choose.

```{r}
bc.integrated <- FindNeighbors(bc.integrated, dims = use.pcs, reduction = "pca")
bc.integrated <- FindClusters(bc.integrated, resolution = seq(0.1,1,0.1), random.seed = 13)
```

Lets first investigate how many clusters each resolution produces 

```{r}
sapply(grep("^RNA_snn_res",colnames(bc.integrated@meta.data),value = TRUE),
       function(x) length(unique(bc.integrated@meta.data[,x])))
```

## Select resolution that best fits your expectations

for the paper of bernard Thienpont they selected 0.5 as their resolution, so here we will go for 0.6

```{r}
Idents(object = bc.integrated) <- "RNA_snn_res.0.6"
```

## Create table displaying the number of each sample belonging to each table

```{r}
table(bc.integrated@active.ident,bc.integrated@meta.data$Sample)
```

## Save Seurat Object

```{r}
save(bc.integrated,file="Data/GandT/RNA/SCEsets/Seurat.RData")
```

## Sesion information

```{r}
sessionInfo()
```

