---
title: "10b_sceG&T_seuratUMAP"
author: "Sebastiaan Vanuytven"
date: "28/08/2019"
output: html_document
---

```{r "setup", include=FALSE}
### Ignore code, this is used to set the workig directory. 
rm(list=ls())
t=getwd()
tt=gsub('.{30}$', '', t)
require("knitr")
opts_knit$set(root.dir = tt)
rm(t,tt)
```

## Load necessary libraries for the analysis

```{r}
library(Seurat)
library(dplyr)
library(viridis)
```

## Parameters to specify

```{r}
use.pcs = 1:7
Colourblind_palette <- c( "#E69F00" , "#56B4E9" , "#009E73" , "#F0E442" , "#0072B2" , "#D55E00" , "#CC79A7" , "#000000" )
```

## load Seurat object

```{r}
load(file="Data/GandT/RNA/SCEsets/Seurat.RData")
```

## Run UMAP

UMAP dimensionality reduction plots are then used to visualise clustering results. As input to the UMAP, you should use the same PCs as input to the clustering analysis.

```{r}
bc.integrated <- RunUMAP(
  object = bc.integrated,
  reduction = "pca",
  dims = use.pcs,
  verbose =  F)
```

## Vizualize UMAP

```{bash}
cd Results/
mkdir -p GandT/RNA/UMAP
```

### Standard tSNE plot coloured by cluster

```{r}
DimPlot(object = bc.integrated, pt.size=3 , label = T , label.size = 12 , reduction = "umap")
ggsave( "Results/GandT/RNA/UMAP/UMAP_standCol.png" , width = 10 , height = 5 )
DimPlot(object = bc.integrated, pt.size=3 , cols = viridis(length(unique(bc.integrated@active.ident))) , label = T , label.size = 12 , reduction = "umap")
ggsave( "Results/GandT/RNA/UMAP/UMAP_ViridisCol.png" , width = 10 , height = 5 )
```

### Standard tSNE plot coloured by Sample origin

```{r}
DimPlot(object = bc.integrated, pt.size=3 , cols = Colourblind_palette , group.by="Sample" , reduction = "umap" )
ggsave( "Results/GandT/RNA/UMAP/UMAP_SampleOrigin.png" , width = 10 , height = 5 )

DimPlot(object = bc.integrated, pt.size=3 , cols = Colourblind_palette , group.by="Sample" , reduction = "umap" )
ggsave( "Results/GandT/RNA/UMAP/UMAP_SampleOrigin.png" , width = 10 , height = 5 )
```

### Standard tSNE plot coloured by cluster and shaped by sample origin

```{r}
DimPlot(object = bc.integrated, pt.size=3 , label = T , shape.by = "Sample", label.size = 12, reduction = "umap")
ggsave( "Results/GandT/RNA/UMAPUMAP_standCol_shapeSource.png" , width = 10 , height = 5 )
DimPlot(object = bc.integrated, pt.size=3 , cols = viridis(length(unique(bc.integrated@active.ident))) , label = T , shape.by = "Sample", label.size = 12, reduction = "umap")
ggsave( "Results/GandT/RNA/UMAP/UMAP_ViridisCol_shapeSource.png" , width = 10 , height = 5 )
```

## TSNE plots coloured by each score of PC

```{r}
FeaturePlot(bc.integrated, features = c("PC_1", "PC_2", "PC_3", "PC_4", "PC_5", "PC_6", "PC_7", "PC_8", "PC_9", "PC_10") , combine = T)
ggsave( "Results/GandT/RNA/UMAP/UMAP_PCscores.png" , width = 20 , height = 10 )
```


## Save Seurat Object

```{r}
save(bc.integrated,file="Data/GandT/RNA/SCEsets/Seurat.RData")
```


## Sesion information

```{r}
sessionInfo()
```


