---
title: "01_sceG&T_createSCE"
author: "Sebastiaan Vanuytven"
date: "28/08/2019"
output: html_document
---

```{r "setup", include=FALSE}
### Ignore code, this is used to set the workig directory. 
rm(list=ls())
t=getwd()
tt=gsub('.{18}$', '', t)
require("knitr")
opts_knit$set(root.dir = tt)
rm(t,tt)
```

# Setting Up the work environment

## Specify necessary information 

```{r}
locCountFile <- "Data/GandT/RNA/htseq_counts_all.csv"
locMetaData <- "Data/GandT/RNA/SampleDescrCombined.txt"
organism <- "Human" # Can be Fly/Mouse/Human
ERCCs <- "Yes"      # Can be Yes/No
id_type <- "ensembl"   # Can be ensembl/symbol
```


## Load the necessary libraries
```{r include=FALSE}
library(SingleCellExperiment)
library(scater)
library(ggplot2)
library(tidyverse)
```


## Load the countfile and sampledescr
remove the .count from the countfile, sort the rows of the metaData alphabetically and

```{r}
counts <- read.csv(locCountFile, header=TRUE, row.names=1)
colnames(counts) = gsub(".counts" , "" , colnames(counts))
counts <- counts[ , order(colnames(counts))]

metaData <- read.table(locMetaData, header=TRUE, row.names=1)
rownames(metaData) <- gsub("-", ".", rownames(metaData))
metaData <- metaData[ order(row.names(metaData)), ]
```

# Create the singleCellExpermint (SCE) object 

```{r}
sce <- SingleCellExperiment(
  assays = list(counts = as.matrix(counts)), 
  colData = metaData)
```

## Convert Ensembl ID to genenames if necessary

```{r}
  if (organism == "Human") {
        dataset <- "hsapiens_gene_ensembl"
        symbol <- "hgnc_symbol"
    } else if (organism == "Mouse") {
        dataset <- "mmusculus_gene_ensembl"
        symbol <- "mgi_symbol"
    } else {
        stop("org should be 'human' or 'mouse'")
    }

    if (id_type == "ensembl") {
        id_type <- "ensembl_gene_id"
    } else if (id_type == "symbol") {
        id_type <- symbol
    } else {
        stop("id_type should be 'ensembl' or 'symbol'")
    }

    sce <- scater::getBMFeatureAnnos(
        sce,
        ids = rownames(sce),
        filters = id_type,
        attributes = c(
            "ensembl_gene_id",
            "external_gene_name",
            symbol,
            "chromosome_name",
            "description",
            "gene_biotype",
            "percentage_gene_gc_content"
        ),
        dataset = dataset,
        host = "www.ensembl.org"
    )

    SummarizedExperiment::rowData(sce)$description <- gsub(
        "\\s\\[.*\\]", "", SummarizedExperiment::rowData(sce)$description
    )


```


## Mark mitochondrial genes and ERCCs in the SCE object

```{r}
if (organism == "Human") {
  if (ERCCs == "Yes") {
    is.mito <- grepl("^MT-", rowData(sce)$hgnc_symbol)
    is.spike <- grepl("^ERCC-", rownames(sce)) 
    rowData(sce)[is.spike,"hgnc_symbol"] = rownames(sce)[is.spike]
    rowData(sce)[is.spike,"ensembl_gene_id"] = rownames(sce)[is.spike]
    rowData(sce)[is.spike,"external_gene_name"] = rownames(sce)[is.spike]
    sce <- calculateQCMetrics(sce, feature_controls=list(ERCC=is.spike, Mt=is.mito))
  } else {
    is.mito <- grepl("^MT-", rowData(sce)$hgnc_symbol)
    sce <- calculateQCMetrics(sce, feature_controls=list(Mt=is.mito))
  }
} else if (Organism == "Mouse") {
  if (ERCCs == "Yes") {
    is.mito <- grepl("^mt-", rowData(sce)$hgnc_symbol)
    rowData(sce)[is.spike,"hgnc_symbol"] = rownames(sce)[is.spike]
    rowData(sce)[is.spike,"ensembl_gene_id"] = rownames(sce)[is.spike]
    rowData(sce)[is.spike,"external_gene_name"] = rownames(sce)[is.spike]
    sce <- calculateQCMetrics(sce, feature_controls=list(ERCC=is.spike, Mt=is.mito))
  } else {
    is.mito <- grepl("^mt-", rowData(sce)$hgnc_symbol)
    sce <- calculateQCMetrics(sce, feature_controls=list(Mt=is.mito))
  }
}
```

```{r}
if (id_type == "ensembl_gene_id") {
  rownames( sce )[rowData(sce)$is_feature_control == F  ] <- 
    make.names( rowData( sce[rowData(sce)$is_feature_control == F ,] )$hgnc_symbol , unique = TRUE )
  sce <- sce[!grepl("NA\\.\\.", rownames(sce)),]
}
```


```{r}
names(colData(sce))
```

```{r}
table(colData(sce)$Sample)
```

```{r}
table(colData(sce)$Plate)
```


```{r}
table(colData(sce)$method)
```

# Save SCE object

make SCEsets folder in Data folder of the project

```{bash}
mkdir -p Data/GandT/RNA/SCEsets
```


```{r}
save(sce,file='Data/GandT/RNA/SCEsets/sce')
```
















