---
title: "03_sceG&T_DataExploration"
author: "Sebastiaan Vanuytven"
date: "28/08/2019"
output: html_document
---

```{r "setup", include=FALSE}
### Ignore code, this is used to set the workig directory. 
rm(list=ls())
t=getwd()
tt=gsub('.{18}$', '', t)
require("knitr")
opts_knit$set(root.dir = tt)
rm(t,tt)
```

# Setting Up the work environment


## Specify necessary information 

```{r}
metricOfInterest <- c("Sample", "Amplification", "Nr_cells" , "Plate") # can be one or multiple elements
ERCCs <- "Yes" # Can be Yes or No
madHistorgamsPerMetricOfInterest <- "Yes" # Can be Yes or No
Organism <- "Human" # can be Human or Mouse
Colourblind_palette <- c( "#E69F00" , "#56B4E9" , "#009E73" , "#F0E442" , "#0072B2" , "#D55E00" , "#CC79A7" , "#000000", "Darkgrey" )
```


## Load the necessary libraries
```{r include=FALSE}
library(SingleCellExperiment)
library(scater)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(scran)
source("Scripts/GandT/RNA/Resources/plotting.R")

set.seed(13)
```

## Load SCE object
```{r}
load("Data/GandT/RNA/SCEsets/sce")
```

## Make metedata only object for plotting

```{r}
metaData <- as.data.frame(colData(sce))
```

# Data Exploration

## Make results folder

```{bash}
mkdir -p Results/GandT/RNA/QC/Overview
mkdir -p Results/GandT/RNA/QC/scatters
```

## Fast exploration of QC metrics with beeswarm plots

### Overview plot of 3 or 4 QC metrics

```{r}
plCounts <- ggplot(metaData, aes(y = metaData[,"log10_total_counts"], x= factor(1))) + 
  geom_violin(colour = "gray60", alpha = 0.2, scale = "width", width = 0.8, fill="grey90") +
  geom_jitter(shape=16, position=position_jitter(0.3), size = 3) +  
  theme_pubr() + 
  xlab("") + ylab("Log10(total counts)\n") + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

plGenes <- ggplot(metaData, aes(y = metaData[,"total_features_by_counts"], x= factor(1))) + 
  geom_violin(colour = "gray60", alpha = 0.2, scale = "width", width = 0.8, fill="grey90") +
  geom_jitter(shape=16, position=position_jitter(0.3), size = 3) +
  theme_pubr() + 
  xlab("") + ylab("Expressed genes\n") + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

plMito <- ggplot(metaData, aes(y = metaData[,"pct_counts_Mt"], x= factor(1))) + 
  geom_violin(colour = "gray60", alpha = 0.2, scale = "width", width = 0.8, fill="grey90") +
  geom_jitter(shape=16, position=position_jitter(0.3), size = 3) +
  theme_pubr() + 
  xlab("") + ylab("% MT counts\n") + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

if (ERCCs == "Yes") {
  
plERCC <- ggplot(metaData, aes(y = metaData[,"pct_counts_ERCC"], x= factor(1))) + 
  geom_violin(colour = "gray60", alpha = 0.2, scale = "width", width = 0.8, fill="grey90") +
  geom_jitter(shape=16, position=position_jitter(0.3), size = 3) +
  theme_pubr() + 
  xlab("") + ylab("% ERCC counts\n") + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
  
  q <- ggarrange( plCounts, plGenes, plMito , plERCC , ncol = 2, nrow=2)
  
} else {
  
 q <- ggarrange( plCounts, plGenes, plMito , ncol = 2, nrow=2)
  
}

ggsave(plot = q, "Results/GandT/RNA/QC/Overview/Generaloverview_violin.png", width = 10 , height = 8 )

q


```

### Overview plot of 3 or 4 QC metrics per interest


```{r}
#vectorToPlot <- c(metricOfInterest, "Nr_cells")
vectorToPlot <- metricOfInterest

for (i in vectorToPlot) {
  
  plCounts <- ggplot(metaData, aes(x = metaData[,i] , y = log10_total_counts)) +
    geom_violin(colour = "gray60", alpha = 0.2, scale = "width", width = 0.8, fill="grey90") + 
    geom_jitter(shape=16, position=position_jitter(0.3), size = 3, aes(color = metaData[,i])) + 
    scale_colour_manual(values=Colourblind_palette) +
    xlab("") + ylab("Log10(total counts)\n") + theme_pubr() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

  plGenes <- ggplot(metaData, aes(x = metaData[,i] , y = total_features_by_counts)) +
    geom_violin(colour = "gray60", alpha = 0.2, scale = "width", width = 0.8, fill="grey90") + 
    geom_jitter(shape=16, position=position_jitter(0.3), size = 3, aes(color = metaData[,i])) + 
    scale_colour_manual(values=Colourblind_palette) +
    xlab("") + ylab("Expressed genes\n") + theme_pubr() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

  plMito <- ggplot(metaData, aes(x = metaData[,i] , y = pct_counts_Mt)) +
    geom_violin(colour = "gray60", alpha = 0.2, scale = "width", width = 0.8, fill="grey90") + 
    geom_jitter(shape=16, position=position_jitter(0.3), size = 3, aes(color = metaData[,i])) + 
    scale_colour_manual(values=Colourblind_palette) +
    xlab("") + ylab("% mitochondrial counts\n") + theme_pubr() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

  if (ERCCs == "Yes") {
  
    plERCC <- ggplot(metaData, aes(x = metaData[,i] , y = pct_counts_ERCC)) +
      geom_violin(colour = "gray60", alpha = 0.2, scale = "width", width = 0.8, fill="grey90") + 
      geom_jitter(shape=16, position=position_jitter(0.3), size = 3, aes(color = metaData[,i])) + 
      scale_colour_manual(values=Colourblind_palette) +
      xlab("") + ylab("% ERCC counts\n") + theme_pubr() + theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
  
    q <- ggarrange( plCounts, plGenes, plMito , plERCC , ncol = 2, nrow=2)
  
  } else {
  
  q <- ggarrange( plCounts, plGenes, plMito , ncol = 2, nrow=2)
  
  }

  ggsave(plot = q, paste0("Results/GandT/RNA/QC/Overview/Generaloverview_violin_", i , ".png"), width = 10 , height = 8 )

  q
  
}



```


## Fast Exploration with MAD histograms

### Overview plot of 3 or 4 QC metrics

Distributions by cell. Blue line shows the median and red lines show median absolute deviations (MADs) from the median

```{r}
if (ERCCs == "Yes") {
  q <- ggarrange(
    outlier_histogram(metaData, "log10_total_counts", mads = 1:6) + ylab("") + xlab("Log10(total counts)") ,
    outlier_histogram(metaData, "log10_total_features_by_counts", mads = 1:6) + ylab("") + xlab("Log10(Expressed genes)"),
    outlier_histogram(metaData, "pct_counts_Mt", mads = c(1, 3, 5, 7)) + ylab("") + xlab("% mitochondrial counts"),
    outlier_histogram(metaData, "pct_counts_ERCC", mads = 1:6) + ylab("") + xlab("% ERCC counts"),
    ncol = 2, nrow=2
  )
} else {
  q <- ggarrange(
    outlier_histogram(metaData, "log10_total_counts", mads = 1:6) + ylab("") + xlab("Log10(total counts)") ,
    outlier_histogram(metaData, "log10_total_features_by_counts", mads = 1:6) + ylab("") + xlab("Log10(Expressed genes)"),
    outlier_histogram(metaData, "pct_counts_Mt", mads = c(1, 3, 5, 7)) + ylab("") + xlab("% mitochondrial counts"),
    ncol = 2, nrow=2
  )
}
ggsave(plot = q, "Results/GandT/RNA/QC/Overview/Generaloverview_mads.png", width = 10 , height = 8 )
q
```


## scatter plots for 4 quality metrics coloured by interest

Results/Pilot/RNA/QC/scatters

```{r}
#vectorToPlot <- c(metricOfInterest, "Nr_cells")

for (i in vectorToPlot) {
  
  pl <- ggplot(metaData, aes(x = log10_total_counts, y = total_features_by_counts)) +
    geom_point(aes(colour = metaData[,i]), size=5) +
    xlab("\n Log10(total counts)") + ylab("Expressed Genes\n") + theme_pubr() +
    labs(color=i) + scale_colour_manual(values=Colourblind_palette)
  ggsave(plot = pl, paste0("Results/GandT/RNA/QC/scatters/Scatter_countsGenes_", i , ".png"), width = 8 , height = 5 )
  
  pl <- ggplot(metaData, aes(x = log10_total_counts, y = pct_counts_Mt)) +
    geom_point(aes(colour = metaData[,i]), size=5) +
    xlab("\n Log10(total counts)") + ylab("\n% mitochondrial counts") + theme_pubr() +
    labs(color=i) + scale_colour_manual(values=Colourblind_palette)
  ggsave(plot = pl, paste0("Results/GandT/RNA/QC/scatters/Scatter_countsMt_", i , ".png"), width = 8 , height = 5 )
  
  if (ERCCs == "Yes") {
    
  pl <- ggplot(metaData, aes(x = pct_counts_Mt, y = pct_counts_ERCC)) +
    geom_point(aes(colour = metaData[,i]), size=5) +
    xlab("\n% mitochondrial counts") + ylab("% ERCC counts\n") + theme_pubr() +
    labs(color=i) + scale_colour_manual(values=Colourblind_palette)
  ggsave(plot = pl, paste0("Results/GandT/RNA/QC/scatters/Scatter_MtERCCs_", i , ".png"), width = 8 , height = 5 )
  
  pl <- ggplot(metaData, aes(x = log10_total_counts, y = pct_counts_ERCC)) +
    geom_point(aes(colour = metaData[,i]), size=5) +
    xlab("\n Log10(total counts)") + ylab("% ERCC counts\n") + theme_pubr() +
    labs(color=i) + scale_colour_manual(values=Colourblind_palette)
  ggsave(plot = pl, paste0("Results/GandT/RNA/QC/scatters/Scatter_countsERCCs_", i , ".png"), width = 8 , height = 5 )
  
  }
} 
```


# Cell Cycle analysis

## Add ensemble ID to rowdat

```{r}
if (Organism == "Human") {
  library(org.Hs.eg.db)
  hm.pairs <- readRDS(system.file("exdata", "human_cycle_markers.rds", package="scran"))
  symb <- mapIds(org.Hs.eg.db, keys=rownames(sce), keytype="SYMBOL", column="ENSEMBL")
  rowData(sce)$SYMBOL <- rownames(sce)
  rowData(sce)$ENSEMBL <- symb
  assignments <- cyclone(sce, hm.pairs, gene.names=rowData(sce)$ENSEMBL)
  sce$CellCycle <- assignments$phases
 } else if (Organism == "Mouse") {
  library(org.Mm.eg.db)
  mm.pairs <- readRDS(system.file("exdata", "mouse_cycle_markers.rds", package="scran"))
  symb <- mapIds(org.Mm.eg.db, keys=rownames(sce), keytype="SYMBOL", column="ENSEMBL")
  rowData(sce)$SYMBOL <- rownames(sce)
  rowData(sce)$ENSEMBL <- symb
  assignments <- cyclone(sce, hm.pairs, gene.names=rowData(sce)$ENSEMBL)
  sce$CellCycle <- assignments$phases
}

```

## Table of cell cycle assignments 

```{r}
kable(table(Phase = colData(sce)$CellCycle, useNA = "ifany"))
```

# Identify High Expression genes

```{r}
plotHighestExprs(sce)
ggsave("Results/GandT/RNA/QC/HighlyExpressedGenes.png", width = 8 , height = 5 )
```

# Quick look at the potential confounding factors

run normailzation and dimensionality reduction on the complete dataset 

```{r}
sizeFactors(sce) <- librarySizeFactors(sce)
sce <- normalize(sce)
sce <- runPCA(sce)
sce <- runTSNE(sce)
```

## Explanatory variables

This plot shows the percentage of variance in the dataset that is explained by various technical factors.

```{r}
#exp_vars <- c("CellCycle", "log10_total_counts",
#              "pct_counts_ERCC", "total_features_by_counts",
#              "pct_counts_Mt","Nr_cells", metricOfInterest)

exp_vars <- c("CellCycle", "log10_total_counts",
              "pct_counts_ERCC", "total_features_by_counts",
              "pct_counts_Mt", metricOfInterest)
all_zero <- Matrix::rowSums(counts(sce)) == 0
plotExplanatoryVariables(sce[!all_zero, ], variables = exp_vars)
ggsave("Results/GandT/RNA/QC/Density_confounders.png", width = 10 , height = 8 )

```

## Dimensionality reduction

Dimensionality reduction plots coloured by technical factors can help identify which may be playing a bit role in the dataset.

```{r}
dimred_factors <- c(
    "Total counts"        = "log10_total_counts",
    "Cell Cycle"           = "CellCycle", 
    "Total features"      = "total_features_by_counts",
    "Mitochondrial genes" = "pct_counts_Mt",
    "ERCCs"                = "pct_counts_ERCC",
    "Number of Cells"     = "Nr_cells"
)
```

```{r warning=FALSE}
plot_list <- lapply(names(dimred_factors), function(fct_name) {
    plotPCA(sce, colour_by = dimred_factors[fct_name] ) +
        ggtitle(fct_name) +
        theme(legend.position = "bottom")
})
cowplot::plot_grid(plotlist = plot_list, ncol = 3)
ggsave("Results/GandT/RNA/QC/PCA_confounders.png", width = 10 , height = 8 )
```

```{r warning=FALSE}
plot_list <- lapply(names(dimred_factors), function(fct_name) {
    plotTSNE(sce, colour_by = dimred_factors[fct_name]) +
        ggtitle(fct_name) +
        theme(legend.position = "bottom")
})
cowplot::plot_grid(plotlist = plot_list, ncol = 3)
ggsave("Results/GandT/RNA/QC/tSNE_confounders.png", width = 10 , height = 8 )
```

```{r}
save(sce,file='Data/GandT/RNA/SCEsets/sce')
```

