---
title: "04_sceG&T_QC"
author: "Sebastiaan Vanuytven"
date: "28/08/2019"
output: html_document
---

```{r "setup", include=FALSE}
### Ignore code, this is used to set the workig directory. 
rm(list=ls())
t=getwd()
tt=gsub('.{18}$', '', t)
require("knitr")
opts_knit$set(root.dir = tt)
rm(t,tt)
```

# Setting Up the work environment


## Specify necessary information 

```{r}
Method <- "Manual" # can be Manual or MAD
ERCCs <- "Yes" # Can be Yes or No
Filtering <- "Manual" # Can be Manual or 10%

## If manual method is preferred for cell filtering 
cutoffCounts  <- 30000 # Advised number for SS2: 100.000
cutoffGenes   <- 500   # Advised number for SS2: > 1000
cutoffMt      <- 20     # Advised number for SS2: < 20-25%
cutoffERCC    <- 10     # Advised number for SS2: < 10%

## If manual method is preferred for gene filtering
min_cells <- 5

### Only to override when certain (consult your local bioinformatician)
nrMAD <- 3 # only to be assigned if using the MAD methodology

Colourblind_palette <- c( "#E69F00" , "#56B4E9" , "#009E73" , "#F0E442" , "#0072B2" , "#D55E00" , "#CC79A7" , "#000000" )
```


## Load the necessary libraries
```{r include=FALSE}
library(SingleCellExperiment)
library(scater)
library(ggplot2)
library(tidyverse)
library(edgeR)
source("Scripts/GandT/RNA/Resources/plotting.R")
set.seed(13)
```

## Load SCE object
```{r}
load("Data/GandT/RNA/SCEsets/sce")
col_data <- as.data.frame(colData(sce))
row_data <- as.data.frame(rowData(sce))
```

# QCing of cells to analyze

## MANUAL 

```{r}

if (Method == "Manual") {
  counts_out   <- sce$total_counts<(cutoffCounts-1)
  features_out <- sce$total_features_by_counts < (cutoffGenes-1)
  mt_out       <- sce$pct_counts_Mt > (cutoffMt)

  kept <- !(counts_out | features_out | mt_out)
  colData(sce)$Kept <- kept
  
  kable(tibble(
    Type = c(
        "Total counts",
        "Total features",
        "Mitochondrial %",
        "Kept cells"
    ),
    Threshold = c(
        paste("< ", cutoffCounts),
        paste("< ", cutoffGenes),
        paste("> ", cutoffMt),
        ""
    ),
    Count = c(
        sum(counts_out),
        sum(features_out),
        sum(mt_out),
        sum(kept)
    )
))
  
  if (ERCCs == "Yes") {
  ERCC_out       <- sce$pct_counts_ERCC > (cutoffERCC)
  
  kept <- !(counts_out | features_out | mt_out | ERCC_out)
  colData(sce)$Kept <- kept
  
  kable(tibble(
    Type = c(
        "Total counts",
        "Total features",
        "Mitochondrial %",
        "ERCC %",
        "Kept cells"
    ),
    Threshold = c(
        paste("< ", cutoffCounts),
        paste("< ", cutoffGenes),
        paste("> ", cutoffMt, "%"),
        paste("> ", cutoffERCC, "%"),
        ""
    ),
    Count = c(
        sum(counts_out),
        sum(features_out),
        sum(mt_out),
        sum(ERCC_out),
        sum(kept)
    )
))
  }
}

```


## MAD

```{r}
if (Method == "MAD") {
  counts_out   <- isOutlier(col_data$log10_total_counts, nmads = 3, type = "lower")
  features_out <- isOutlier(col_data$log10_total_features_by_counts, nmads = 3, type = "lower")
  mt_out       <- isOutlier(col_data$pct_counts_Mt, nmads = 3, type = "higher")
  
  counts_thresh   <- attr(counts_out, "thresholds")["lower"]
  features_thresh <- attr(features_out, "thresholds")["lower"]
  mt_thresh       <- attr(mt_out, "thresholds")["higher"]

  kept <- !(counts_out | features_out | mt_out)
  colData(sce)$Kept <- kept
  
  kable(tibble(
    Type = c(
        "Total counts",
        "Total features",
        "Mitochondrial %",
        "Kept cells"
    ),
    Threshold = c(
        paste("< 10 ^", round(counts_thresh, 2),
              paste0("(", round(10 ^ counts_thresh), ")")),
        paste("< 10 ^", round(features_thresh, 2),
              paste0("(", round(10 ^ features_thresh), ")")),
        paste(">", round(mt_thresh, 2), "%"),
        ""
    ),
    Count = c(
        sum(counts_out),
        sum(features_out),
        sum(mt_out),
        sum(kept)
    )
))
  
  if (ERCCs == "Yes") {
  ERCC_out       <- isOutlier(col_data$pct_counts_ERCC, nmads = 3, type = "higher")
  ERCC_thresh    <- attr(ERCC_out, "thresholds")["higher"]
  
  kept <- !(counts_out | features_out | mt_out | ERCC_out)
  colData(sce)$Kept <- kept
  
  kable(tibble(
    Type = c(
        "Total counts",
        "Total features",
        "Mitochondrial %",
        "ERCC %",
        "Kept cells"
    ),
    Threshold = c(
        paste("< 10 ^", round(counts_thresh, 2),
              paste0("(", round(10 ^ counts_thresh), ")")),
        paste("< 10 ^", round(features_thresh, 2),
              paste0("(", round(10 ^ features_thresh), ")")),
        paste(">", round(mt_thresh, 2), "%"),
        paste(">", round(ERCC_thresh, 2), "%"),
        ""
    ),
    Count = c(
        sum(counts_out),
        sum(features_out),
        sum(mt_out),
        sum(ERCC_out),
        sum(kept)
    )
))
  }
}
```

## Filter out bad quality cells

```{r}
sce_qc <- sce[, kept]
```


# Gene Filtering

```{r}
if (Filtering == "Manual") {
  keep <- Matrix::rowSums(counts(sce_qc) >= 1) >= min_cells
  sce_qc <- sce_qc[keep, ]
} else if (Filtering == "10%") {
  min_cells <- ceiling(ncol(sce)/10)
  keep <- Matrix::rowSums(counts(sce_qc) >= 1) >= min_cells
  sce_qc <- sce_qc[keep, ]
}
  
```

# Sanity check if we removed an interesting population with our QCíng

One thing we can look at is the difference in expression between the kept and removed cells. If we see known genes that are highly expressed in the removed cells that can indicate that we have removed an interesting population of cells from the dataset. The red line shows equal expression and the blue line is a linear fit.

```{r}
pass_qc <- colnames(sce) %in% colnames(sce_qc)
lost_counts <- counts(sce)[, !pass_qc]
kept_counts <- counts(sce)[,  pass_qc]

kept_lost <- tibble(
    Gene     = rownames(sce),
    Lost     = calcAverage(lost_counts),
    LostProp = Matrix::rowSums(lost_counts > 0) / ncol(lost_counts),
    Kept     = calcAverage(kept_counts),
    KeptProp = Matrix::rowSums(kept_counts > 0) / ncol(kept_counts)
) %>%
    dplyr::mutate(LogFC = predFC(cbind(Lost, Kept),
                          design = cbind(1, c(1, 0)))[, 2]) %>%
    dplyr::mutate(LostCapped = pmax(Lost, min(Lost[Lost > 0]) * 0.5),
           KeptCapped = pmax(Kept, min(Kept[Kept > 0]) * 0.5))

ggplot(kept_lost,
       aes(x = LostCapped, y = KeptCapped, colour = LostProp - KeptProp)) +
    geom_point(size = 1, alpha = 0.2) +
    geom_abline(intercept = 0, slope = 1, colour = "red") +
    geom_smooth(method = "lm") +
    scale_x_log10() +
    scale_y_log10() +
    scale_colour_viridis_c() +
    ggtitle("Gene expression in kept and removed cells") +
    xlab("Average count (removed)") +
    ylab("Average count (kept)") +
    theme_minimal()
```

```{r}
kept_lost %>%
    dplyr::select(Gene, LogFC, Lost, LostProp, Kept, KeptProp) %>%
    dplyr::arrange(-LogFC) %>%
    as.data.frame() %>%
    head(10)
```

# removing the remaining controls

```{r}
#table(colData(sce_qc)$Nr_cells)
#sce_qc <- sce_qc[,colData(sce_qc)$Nr_cells == "SC"  ]

table(colData(sce_qc)$SCMCNC)
sce_qc <- sce_qc[,colData(sce_qc)$Nr_cells == "SC"  ]
```


# Save SCE object after quality control

```{r}
save(sce_qc,file='Data/GandT/RNA/SCEsets/sce_qc')
```
